---
title: PHYLLOTAXIS - Draw Flowers using Math
output: github_document
---
# Patterns in Nature 
Thereare many examples of natural facts that can be described in mathematical terms. Nice examples are the shape of snowflakes, the fractal geometry of romanesco broccoli or how self-similarity rules the growth of plants.

R is a tool for doing serious analysis, but not everything in life is serious. Life is also funny, and R can be used to have fun and to do beautiful things. Its graphical power can be used to produce artistic images like the one that illustrates this section, which is inspired by how plants arrange their leaves. This fact is called phyllotaxis and will serve as the basis of this project.

```{r}
library(ggplot2)
```

There are many ways to represent data with ggplot2: from simple scatter plots to more complex violin plots. The functions that start with geom_ define the type of plot. In this notebook, we will only work with geom_point() which plots points in two dimensions. We'll need a dataset with two variables; let's call them x and y.

We'll start by drawing 50 points on a circle of radius 1. As every (x, y) point should be in the unit circle, it follows that x² + y² = 1. We can get this using the super famous Pythagorean trigonometric identity which states that sin²(θ) + cos²(θ) = 1 for any real number θ.
```{r}
# Create circle data to plot
t <- seq(0, 2*pi, length.out = 50)
x <- sin(t)
y <- cos(t)
df <- data.frame(t, x, y)

# Make a scatter plot of points in a circle
p <- ggplot(df, aes(x, y))
p + geom_point()

```

## Create a scatter plot of spiralized plots

```{r}
# Define the number of points
points <- 500

# Define the Golden Angle

angle <- pi * (3 - sqrt(5))

t <- (1:points) * angle
x <- sin(t)
y <-cos(t)
df <- data.frame(t, x, y)

# Make a scatter plot of points in a spiral
p <- ggplot(df, aes(x*t, y*t))
p + geom_point()
```
Remove everything unnecessary 

```{r}
df <- data.frame(t, x, y)

# Make a scatter plot of points in a spiral and remove some plot components
p <- ggplot(df, aes(x*t, y*t))
p + geom_point() +
# .... YOUR CODE FOR TASK 4 ....
 theme(panel.background = element_rect(fill="white"),
        axis.text=element_blank(), axis.ticks = element_blank())

```

```{r}
# Change the code from Task 4 to modify the 
# size, transparency, and color of the points
p <- ggplot(df, aes(x*t, y*t))
p + geom_point(size = 8, alpha = 0.5, color = "darkgreen") 
```

```{r}
# Copy the code from Task 5 and modify the 
# color, size, and shape of the points
p <- ggplot(df, aes(x*t, y*t))
p +  geom_point (aes(size=t),shape = 8 )
theme(legend.position = "none")
```

## Put it all together: the sunflower 


```{r}
p <- ggplot(df, aes(x*t, y*t))
p + geom_point(aes(size=t), alpha=0.5, shape=17, color="yellow") +
    theme(legend.position="none",
          panel.background = element_rect(fill="darkmagenta"))
```

