---
title: "Structural Bioinormatics Part 1."
author: "Vicky Vu"
date: "2/13/2020"
output: github_document
---

---
title: "Class_12: Structural Bioinformatics pt. 1"
author: "Vicky Vu"
date: "2/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB Statistics 

Here we inspect the types of structures in themain database for 3D biomolecular data - the PDB. 

Q1: Determine the percentage of structures solved by X-Ray and Electron Microscopy. Also can
you determine what proportion of structures are protein? 
```{r}
#read in the file 
stats <- read.csv("Data Export Summary.csv", row.names = 1)
stats

#calculate percentage values asked for  
sum  <- sum (stats [1,5],stats[3,5])
total <- sum ( stats [,5])

sum/total *100

```

```{r}
# percentage of each 
stats$Total/sum(stats$Total)*100

# what porportion of structures are protein 
sum(stats$Proteins)/sum(stats$Total)*100
```



Q2: Type HIV in the PDB website search box on the home page and determine how many
HIV-1 protease structures are in the current PDB? 1289 HIV-1 protease structures found in te current PDB.

# part 2 
Read a single PDB structure into R 
```{r}
library(bio3d)

pdb <- read.pdb("1hsg")
pdb
```
# Select protein only 

Heere I will use the 'atom.select()' function to do this...

```{r}
lig <- atom.select(pdb, "ligand", value = TRUE)

prot <- atom.select(pdb, "protein", value = TRUE)
```
...and write out these new PDB objects with the "write.pdb()' function. 
```{r}
write.pdb(lig, file = "1hsg_ligand.pdb")
write.pdb( prot, file = "1hsg_protein.pdb")
```


```{r}
attributes(pdb)
```







