---
title: "Class 13: Structure based drug discovery"
author: "Vicky Vu"
date: "2/18/2020"
output: github_document
---

## Download and clean HIV-Pr receptor structure 

Here we will download the 1HSG structure from the online database,
remove the bound drug from Merk and write out a protein only and ligand only new set of PDB files.

We wil use the functions 'get.pdb()', 'read.pdb()', atom.select()', and finally 'write.pdb()'. 

```{r}
#bio3d was already downloaded 
library(bio3d)

# Download PDB file 
file.name <- get.pdb ("1hsg")
```

Here we'll prepare protein and ligand input files.
```{r}
# Here you can get a quick summary of the pdb structure object 
pdb<- read.pdb(file.name)
```

Select and write protein only 
```{r}
prot <- atom.select (pdb,"protein",value=TRUE )
write.pdb (prot, file="1hsg_protein.pdb")

```

select and write ligand only
```{r}
lig <- atom.select(pdb, "ligand", value=TRUE)
write.pdb (lig, file = "1hsg_ligand.pdb")
```

Using AutoDock Tools to setup protein docking input 
AutoDock Tools is found in MGLTools. 

In ADT (AutoDocTools), load the protein using File > Read Molecule.
Select 1hsg_protein.pdb. Click Open


Next prepare a docking configuration file

## Section 2: Docking ligands into HIV-1 protease 

Read docking results 
```{r}
res <- read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb")

```
















