---
title: 'Class 6: R Function'
author: "Vicky Vu"
date: "1/23/2020"
output:
  word_document: default 
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions 

Read table functions 
  1. read.table("filename.txt", header=TRUE, sep="\t")
  
  2. separater type consists of \t  "," is comma, ";" is semicolin, "" is white space 
  
  3. read.delim
  
Lets see more about **file import** (i.e. reading files into R). The main read function in base R is `read.table()`

```{r}
t1 <- read.table("test1.txt", header=TRUE, sep=",")
t1
```
You could also use `read.csv` which has the arguments I want in this case.
```{r}
t1 <- read.csv("test1.txt")
t1
```


```{r}
t2 <- read.table("test2.txt", header=TRUE, sep="$")
t2
```


```{r}
t3 <- read.table("test3.txt")
t3
```

# Back To Functions
```{r}
add <- function(x, y=1) {
 # Sum the input x and y
 x + y
} 
```

using this function 

```{r}

add(7,3)
add(x=1, y=4)
add(1, 4)
add(1)
add(3)

```

How does this work with ventors?
```{r}
add( c(1, 2, 4) )
add( c(1, 2, 4), 4 )
add( c(1, 2, 4), c(4, 5, 6) )

```


What is this `range()` function?
```{r}
x <- c(4,4,10,11,1)
max(x)
min (x)
range(x)
rng <- range(x)
rng
```
This is our second function 



```{r}
rescale <- function(x) {
 rng <-range(x)
 (x - rng[1]) / (rng[2] - rng[1])
}
rescale(1:10)
rescale( c(1,2,NA,3,10) )

```
```{r}
rescale2 <- function(x) {
  rng <- range (x, na.rm=TRUE)
   (x - rng[1]) / (rng[2] - rng[1])
}

```

```{r}
rescale2( c(1,2,NA,3,10) )
```


Going too far!
```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
 rng <-range(x, na.rm=na.rm)
 print("Hello")
 
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 
 print("is it me you are looking for?")
 
 if(plot) {
   print("don't sing again")
 plot(answer, typ="b", lwd=4)
 }
 print("I can see it in ...")
 return(answer)
}
```
```{r}
rescale3(x, plot=TRUE)
```

```{r}
rescale3(x, plot=FALSE)
```

## Hands on Worksheet Part B 
```{r}

library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")

```
Q1. What type of object is returned from the read.pdb() function?
It is a list of 8 things and of class "pdb", "sse"
```{r}
s1

```
```{r}
class(s1)

```
```{r}
str(s1)
```
```{r}
aa321(s1$seqres)
```

```{r}
s1$atom
```


Q2. What does the trim.pdb() function do?
```{r}
trim.pdb(s1, chain="A", elety="CA")
```


It makes it smaller. 

```


Q3. What input parameter would turn off the marginal black and grey rectangles in the
plots and what do they represent in this case?


Q4. What would be a better plot to compare across the different proteins?
If the 3 plots where put together into 1 plot. 




```

```




Q5. Which proteins are more similar to each other in their B-factor trends. How could
you quantify this?
s1.b and s3.b are more similar to each other 
```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)

```




Q6. How would you generalize the original code above to work with any set of input
protein structures?




### Level 3 Heading 












